---
subtitle: "**SIOFA**"
title: "00:"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
#knit: (function(inputFile, encoding) {
#  rmarkdown::render(inputFile, encoding = encoding, output_dir="html")})
bibliography: refs.bib
editor_options: 
  markdown: 
    wrap: 72
---


[Jump to More Information](#more)

[Jump to References](#references)


```{r, knitr, echo=FALSE, warning=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =TRUE, 
               eval      =TRUE,
               
               cache     =TRUE, 
               cache.path="cache/00/",
               
               fig.path  ="../tex/figs/00-",
               fig.width =8,
               fig.height=6,
               dev       ="png")

options(digits=3)

iFig=0
iTab=0
```

## Introduction {#intro}



## Libraries

A variety of R packages are required for data manipulation and plotting,
obtaining life history parameters, and modelling.

#### Data Manipulation 

```{r, pkgs-data}
library(plyr)
library(dplyr)
library(reshape)
```

#### Plotting

```{r, pkgs-ggplot}
library(ggplot2)
library(ggpubr)
library(ggcorrplot)
library(GGally)

library(scales)

library(R.devices)
```

#### FLR

```{r, pkgs-flr}
library(FLCore)
library(ggplotFL)
library(FLBRP)
library(FLasher)
library(FLife)
library(mydas)
```


There are also extensive tutorials. 

#### Life Histories

```{r, pkgs-fb}
library(rfishbase)
library(FishLife)
library(SPMpriors)
```

#### Utilities

```{r, pkgs-fb}
library(statcomp)
```

### Installation 

First the required Libraries (i.e. packages) need to be installed. This can be be done
from cran, e.g.

Set mirror if needed `chooseCRANmirror()`

```{r, install-util, eval=FALSE}
install.packages(c("plyr", 
                   "dplyr", 
                   "reshape"))
```

```{r, install-gg, eval=FALSE}
install.packages(c("ggplot2",
                   "ggpubr",
                   "ggcorrplot",
                   "GGally"))
```

#### FLR

Or from <https://flr-project.org/>

To install the latest released versions of the FLR packages, and their dependencies
enter

```{r, eval=FALSE}
source("http://flr-project.org/R/instFLR.R")
```

then select 'FLCore', 'ggplotFL', 'FLBRP', 'FLasher', 'FLife', 'mydas'

Or install from source

```{r, eval=FALSE}
chooseCRANmirror()

install.packages("remotes")

library(remotes)

remotes:::install_github("flr/FLCore")
remotes:::install_github("flr/FLBRP")
remotes:::install_github("flr/ggplotFL")
remotes:::install_github("flr/FLfishery")
remotes:::install_github("flr/FLasher", dependencies=TRUE)
remotes:::install_github("flr/FLife")
remotes:::install_github("flr/mydas")
```


#### FishBase

There are compatability problems with `FishBase`, so install an earlier
version. 

These have been fixed now. see <https://james-thorson-noaa.github.io/FishLife/>

```{r, eval=FALSE}
#remotes::install_github("ropensci/rfishbase",ref="d150f2e0f5")
remotes::install_github("james-thorson/FishLife")
remotes::install_github("Henning-Winker/SPMpriors")
```


[Back to Top](#top)

## Author information

**Laurence Kell**.
[laurie\@seaplusplus.co.uk](mailto:laurie@seaplusplus.co.uk){.email}

## Acknowledgements

## Software Versions

**`r version$version.string`**

-   knitr: `r packageVersion('knitr')`

-   FLCore: `r packageVersion('FLCore')`

-   FLBRP: `r packageVersion('FLBRP')`

-   FLasher: `r packageVersion('FLasher')`

-   FLife: `r packageVersion('FLife')`

-   ggplotFL: `r packageVersion('ggplotFL')`

-   rfishbase: `r packageVersion('rfishbase')`

-   SPMpriors: `r packageVersion('SPMpriors')`

-   FishLife: `r packageVersion('FishLife')`

-   plyr: `r packageVersion('plyr')`

-   dplyr: `r packageVersion('dplyr')`

-   reshape: `r packageVersion('reshape')`

-   ggplot2: `r packageVersion('ggplot2')`

-   ggpubr: `r packageVersion('ggpubr')`

-   ggcorrplot: `r packageVersion('ggcorrplot')`

-   GGally: `r packageVersion('GGally')`

**Compiled**: `r date()`

[Back to Top](#top)

[Back to Top](#top)

# References {#references}

[Back to Top](#top)
