---
title: "**Annex: 00**"
subtitle: "Set Up"
author: "DELEGATION OF THE EUROPEAN UNION"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: SIOFA, deepwater sharks, hierchical framework
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
editor_options: 
  markdown: 
    wrap: 72
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir="html")})
---

[Jump to More Information](#more)

[Jump to References](#references)

```{r, knitr, echo=FALSE, warning=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =!TRUE, 
               eval      =TRUE,
               
               cache     =TRUE, 
               cache.path="cache/00/",
               
               fig.path  ="../tex/figs/00/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

options(digits=3)

iFig=0
iTab=0
```

## Libraries {#libs}

A variety of R packages are required for data manipulation and plotting,
obtaining life history parameters, and modelling.

### Data Manipulation

```{r}
library(plyr)       # Set of tools for splitting, applying, and combining data.
library(dplyr)      # Grammar of data manipulation, a set of verbs that solves the most common data manipulation challenges.
library(reshape)    # Reshaping and aggregating data
```

### Plotting

```{r}
library(ggplot2)    # Grammar of Graphics.
library(ggpubr)     # Wrapper for creating publication-ready plots using ggplot2.
library(ggcorrplot) # Visualisation of correlation matrices using 'ggplot2'.
library(GGally)     # Etension to 'ggplot2', providing additional plotting

library(scales)     # Scaling for 'ggplot2' normalisation and mapping.

library(R.devices)  # Functions for handling graphical output devices.
```

### FLR (Fisheries Library in R)

```{r}
library(FLCore)     # Core FLR package 
library(ggplotFL)   # Extends 'ggplot2' for common fisheries plots using FLR objects.
library(FLBRP)      # Rererence points.
library(FLasher)    # Short-term forecasts and MSE.
library(FLife)      # Life history analysis
library(mydas)      # Data-limited management strategy evaluation.
```

### Life Histories

```{r}
library(rfishbase)  # Interface to 'FishBase'.
library(FishLife)   # Life history parameters for fish species, built on 'FishBase'.
library(SPMpriors)  # Generates prior distributions for stock productivity models.
```

### Utilities

```{r}
library(openxlsx)     # Reading, writing, and editing Excel files

library(lubridate)    # Makes it easier to work with dates and times.

library(statcomp)     # Tools for statistical complexity and entropy based analysis of time series data.

library(randomForest) # Breiman's Random Forest algorithm for classification and regression.

library(mgcv)         # GAMs

library(ggstats)      # Weighted quantiles

library(kableExtra)   # Markdown and Latex tables
```

```{r}
library(fishmethods) #powertrends
```


```{r, source}
source("../source/fl_mcupd_fn.R")
source("../source/fnDat.R") 
```

### Installation

First the required Libraries (i.e. packages) need to be installed. This
can be be done from cran, e.g.

Set mirror if needed `chooseCRANmirror()`

```{r, install-util, eval=FALSE}
install.packages("randomForest")

install.packages(c("plyr", 
                   "dplyr", 
                   "reshape"))

install.packages(c("ggplot2",
                   "ggpubr",
                   "ggcorrplot",
                   "GGally"))
```

#### FLR

Or from <https://flr-project.org/>

To install the latest released versions of the FLR packages, and their
dependencies enter

then select 'FLCore', 'ggplotFL', 'FLBRP', 'FLasher', 'FLife', 'mydas'

Or install from source

```{r, eval=FALSE}
chooseCRANmirror()

install.packages("remotes")

library(remotes)

remotes:::install_github("flr/FLCore")
remotes:::install_github("flr/FLBRP")
remotes:::install_github("flr/ggplotFL")
remotes:::install_github("flr/FLfishery")
remotes:::install_github("flr/FLasher", dependencies=TRUE)
remotes:::install_github("flr/FLife")
remotes:::install_github("flr/mydas")
```

#### FishBase

There are compatability problems with `FishBase`, so install an earlier
version.

These have been fixed now. see
<https://james-thorson-noaa.github.io/FishLife/>

```{r, eval=FALSE}
#remotes::install_github("ropensci/rfishbase",ref="d150f2e0f5")
install.packages("../rfishbase_3.1.10.tar.gz", repos=NULL, type="source")

remotes::install_github("james-thorson/FishLife")
remotes::install_github("Henning-Winker/SPMpriors")
```

[Back to Top](#top)

## Data

### SIOFA

```{r, data-siofa, echo=FALSE}
# Load data from Excel file into data frames
cpue=read.xlsx("../data/inputs/SIOFA_DWS-2023-01-Data.xlsx", "dws_data_catch", 
                 startRow=1, detectDates=TRUE, skipEmptyRows=TRUE)
names(cpue)[c(3,6)]=tolower(names(cpue)[c(3,6)])
names(cpue)[c(7:8)]=c("lat","lon")
cpue$SettingTime=adjust_time_format(cpue$SettingTime)
cpue$HaulingTime=adjust_time_format(cpue$HaulingTime)
cpue$time       =adjust_time_format(cpue$time)
cpue$year       =year(cpue$Date)
names(cpue)[9]  ="depth"

biol=read.xlsx("../data/inputs/SIOFA_DWS-2023-01-Data.xlsx", "dws_data_bio_sampling",
                 startRow=1, detectDates=TRUE, skipEmptyRows=TRUE)

biol=transform(biol, gender=factor(Sex,labels=c("Female","Male"),levels=c("F","M")))
biol=merge(biol,cpue[,c("OPE_ID","gear","depth","Date","year","lat","lon")],by="OPE_ID")
names(biol)[11]="len"

cSmry=catchSmry(cpue)

names(biol)[ 3]="species"
names(cpue)[16]="species"
names(cSmry)[16]="species"

siofa=list("biol"=biol,"cpue"=cpue,"cSmry"=cSmry)

save(siofa, file="../data/siofa.RData")
```

### IEO

```{r, data-ieo, echo=FALSE}
load("../data/inputs/Sharks_logbooks.Rdata")
cpue=as.data.frame(cpue)
names(cpue)[13]="species"

load("../data/inputs/Sharks_biological.Rdata")
biol=as.data.frame(biol)
names(biol)[c(3,5,6,7,12)]=c("len","mass","sex","mat","species")

biol=merge(biol,cpue[,c("cod_set","month","year","Area","x3","x4","y3","y4")],by=c("cod_set"))
names(biol)=tolower(names(biol))
biol=transform(biol,lat=(y3+y4)/2,lon=(x3+x4)/2)

mat=setNames(seq(4),c("A","B","C","D"))
biol$mat[biol$mat%in%names(mat)]=mat[biol$mat[biol$mat%in%names(mat)]]
biol$mat=substr(biol$mat,nchar(biol$mat),1)

biol=transform(biol,spp=factor(sppAbbrv(species)))
biol=transform(biol,mature=ifelse(as.numeric(mat)>=3,1,0))

biol$sex[is.na(biol$sex)]="N"
biol=transform(biol, gender=factor(sex,labels=c("Female","Male","Unsexed"),levels=c("F","M","N")))

set.seed(123) # Setting a seed for reproducibility
rf=randomForest(depth~lat+lon, 
                data=subset(siofa$cpue,!(is.na(depth)|is.na(lat)|is.na(lon))))
biol$depth=predict(rf, newdata=subset(biol,!(is.na(lat)|is.na(lon))))

# Check and correct species scientific names for the cpue dataset
biol$species[biol$species == "Centrophrous granulosus"] <- "Centrophorus granulosus"
biol$species[biol$species == "Centrophorus squamulosus"] <- "Centrophorus squamosus"
biol$species[biol$species == "Centrophorus Squamulosus"] <- "Centrophorus squamosus"
biol$species[biol$species == "Centroscymnus Coelolepis"] <- "Centroscymnus coelolepis"
biol$species[biol$species == "Dalatias Licha"] <- "Dalatias licha"
biol$species[biol$species == "Deania Calcea"] <- "Deania calcea"
biol$species[biol$species == "Etmopterus  spp"] <- "Etmopterus spp"
biol$species[biol$species == "Etmopterus sp"] <- "Etmopterus spp"
biol$species[biol$species == "Etmopterus Spp"] <- "Etmopterus spp"
biol$species[biol$species == "Etmopterus Spp"] <- "Etmopterus spp"

# Check and correct species scientific names for the cpue dataset

cpue$species[cpue$species == "Centrophrous granulosus"] <- "Centrophorus granulosus"
cpue$species[cpue$species == "Centrophorus squamulosus"] <- "Centrophorus squamosus"
cpue$species[cpue$species == "Centrophorus Squamulosus"] <- "Centrophorus squamosus"
cpue$species[cpue$species == "Centroscymnus Coelolepis"] <- "Centroscymnus coelolepis"
cpue$species[cpue$species == "Dalatias Licha"] <- "Dalatias licha"
cpue$species[cpue$species == "Deania Calcea"] <- "Deania calcea"
cpue$species[cpue$species == "Etmopterus  spp"] <- "Etmopterus spp"
cpue$species[cpue$species == "Etmopterus sp"] <- "Etmopterus spp"
cpue$species[cpue$species == "Etmopterus Spp"] <- "Etmopterus spp"
cpue$species[cpue$species == "Etmopterus Spp"] <- "Etmopterus spp"

ieo=list("biol"=biol,"cpue"=cpue)

save(ieo,  file="../data/ieo.RData") 
```

## Species List

```{r, spp, echo=FALSE}
asfis=read.csv("../data/inputs/ASFIS_sp_2021.txt")

sppMrag=c(
"Portuguese dogfish"  ,"Centroscymnus coelolepis",	
"Kitefin shark"       ,"Dalatias licha",	
"Birdbeak dogfish"    ,"Deania calcea",												
"Gulper shark"        ,"Centrophorus granulosus",
"Gulper shark"        ,"Centrophorus atromarginatus",	
"Gulper shark"        ,"Centrophorus squamosus",	
"Gulper shark"        ,"Centrophorus harrissoni",	
"Gulper shark"        ,"Centrophorus isodon",	
"Gulper shark"        ,"Centrophorus lesliei",	
"Gulper shark"        ,"Centrophorus lusitanicus",	
"Gulper shark"        ,"Centrophorus moluccensis",	
"Gulper shark"        ,"Centrophorus seychellorum",	
"Gulper shark"        ,"Centrophorus uyato",	
"Gulper shark"        ,"Centrophorus westraliensis",	
"Gulper shark"        ,"Centrophorus zeehaani",	
"Lantern shark"       ,"Etmopterus alphus",	
"Lantern shark"       ,"Etmopterus bigelowi",	
"Lantern shark"       ,"Etmopterus brosei",	
"Lantern shark"       ,"Etmopterus compagnoi",	
"Lantern shark"       ,"Etmopterus evansi",	
"Lantern shark"       ,"Etmopterus fusus",	
"Lantern shark"       ,"Etmopterus gracilispinis",	
"Lantern shark"       ,"Etmopterus granulosus",	
"Lantern shark"       ,"Etmopterus pusillus",	
"Lantern shark"       ,"Etmopterus sentosus")
sppMrag=data.frame(name=sppMrag[seq(1,length(sppMrag),2)],spp=sppMrag[seq(2,length(sppMrag),2)])

sppSiofa=c("Centroscymnus coelolepis","Pailona commun","Portugese dogfish","CYO",
           "Deania calcea","Squale savate","Birdbeak dogfish","DCA",
           "Centrophorus granulosus","Requin chagrin ","ulper shark","GUP",
           "Dalatias licha","Squale liche","Kitefin shark","SCK",
           "Bythaelurus bachi","Requin chat de Bach","Bach's catshark","BZO",
           "Chimaera buccanigella","Chimere bouche-foncee","Dark-mouth chimaera","ZZC",
           "Chimaera didierae","Chimere de Didier","The Falkor chimaera","ZZD",
           "Chimaera willwatchi","Chimere du marin","Seafarer's ghostshark","ZZE",
           "Centroscymnus crepidater","Pailona à long nez","Longnose Velvet Dogfish","CYP",
           "Centroscymnus plunketi","Pailona austral","Plunket shark","CYU",
           "Zameus squamulosus","Squale-grogneur a queue echancree","Velvet dogfish","SSQ",
           "Etmopterus alphus","Requin lanterne à joues blanches","Whitecheek lanternshark","EZU",
           "Apristurus indicus","Holbiche artouca","Smallbelly catshark","APD",
           "Harriotta raleighana","Chimere à nez rigide","Bentnose rabbitfish","HCR",
           "Bythaelurus tenuicephalus","Requin chat à tête étroite","Narrowhead catshark","BZL",
           "Chlamydoselachus anguineus","Requin lézard","Frilled shark","HXC",
           "Hexanchus nakamurai","Requin griset","Bigeyed six-gill shark","HXN",
           "Etmopterus pusillus","Sagre nain","Smooth lanternshark","ETP",
           "Somniosus antarcticus","Requin dormeur antarctique","Southern sleeper `shark","SON",
           "Mitsukurina owstoni","Requin lutin","Goblin shark","LMO")
sppSiofa=as.data.frame(t(array(sppSiofa,c(4,20))))
sppSiofa=cbind(ldply(strsplit(sppSiofa[,1]," ")),sppSiofa)
names(sppSiofa)=c("Genus","Species","spp","french","name","FAO")

inventory=rbind(
      data.frame(species=unique(ieo$biol$species),  Source="IEO",  What="Biol"),
      data.frame(species=unique(ieo$cpue$species),  Source="IEO",  What="CPUE"),
      data.frame(species=unique(siofa$biol$species),Source="SIOFA",What="Biol"),
      data.frame(species=unique(siofa$cpue$species),Source="SIOFA",What="CPUE"))
old=inventory[,1]
new=inventory[,1]

new=tolower(new)
new=gsub("Spp","spp",                  new)
new=gsub("spp","sp",                   new)
new=gsub("sp", "spp",                  new)
new=gsub("  "," ",                     new)
new=gsub("  "," ",                     new)
new=gsub("calceus","calcea",           new)
new=gsub("squamosus","squamulosus",    new)
new=gsub("Centrophrous","Centrophorus",new)

substr(new,1,1)=toupper(substr(new,1,1))

inventory[,1]    =new
inventory        =inventory[!duplicated(inventory),]
inventory$Genus  =laply(strsplit(inventory$species, " "), function(x) x[[1]])
inventory$Species=laply(strsplit(inventory$species, " "), function(x) x[[2]])
names(inventory)[1]="spp"

spp=sort(unique(c(inventory$spp,sppSiofa$spp,sppMrag$spp)))
spp=cbind(spp    =spp,
          Genus  =laply(strsplit(spp, " "), function(x) x[[1]]),
          Species=laply(strsplit(spp, " "), function(x) x[[2]]))
spp=spp[order(spp[,"spp"]),]

sppFB    =transform(rfishbase::fishbase,spp=paste(Genus,Species))

spp      =cbind(spp,       fishbase=spp[,      "spp"]%in%sppFB$spp)
inventory=cbind(inventory, fishbase=inventory[,"spp"]%in%sppFB$spp)

names(asfis)[4]="spp"
asfis=asfis[asfis[,4]%in%spp[,1],]

fb   =popgrowth(spp[,"spp"])[,c("Species","K","Loo","to","LmMale","LmFemale","Lm")]
names(fb)=c("species","k","linf","t0","l50m","l50f","l50")
depth=species(sppSiofa[,3])[c("Species","DepthRangeDeep")]
names(depth)=c("species","range")
fb   =merge(depth,fb)

save(spp,inventory,sppSiofa,sppMrag,asfis,fb,file="../data/spp.RData")
```

```{r, priors, eval=FALSE}
priors=mdply(spp[,c("Genus","Species")], function(Genus,Species) { 
  
  rtn=try(Plot_taxa(Search_species(Genus=Genus,Species=Species)$match_taxonomy,mfrow=c(3,2),plot=FALSE))
  
  if ("try-error"%in%is(rtn)) return(NULL)
  data.frame(level=length(rtn),trait=names(rtn[[1]][[2]]),data=rtn[[1]][[2]],var=diag(rtn[[1]][[1]]))})
prior2=mdply(spp[,c("Genus","Species")], function(Genus,Species) { 
  
  rtn=try(Plot_taxa(Search_species(Genus=Genus)$match_taxonomy,mfrow=c(3,2),plot=FALSE))
  
  if ("try-error"%in%is(rtn)) return(NULL)
  data.frame(level=length(rtn),trait=names(rtn[[1]][[2]]),data=rtn[[1]][[2]],var=diag(rtn[[1]][[1]]))})
priors=rbind(cbind(What="Species",priors),cbind(What="Genus",prior2))
priors=priors[!duplicated(priors[,c("Genus","Species","trait")]),]

traits=mdply(spp[,c("Genus","Species")], function(Genus,Species) {
  
  rtn=try(fl_mcupd(Genus=Genus, Species=Species)$traits)
  
  if ("try-error"%in%is(rtn)) return(NULL)
  rtn})

for (i in list.files(pattern = "\\.png$")){
  file.copy(  i,file.path("../figs/png",i))
  file.remove(i)}

save(priors,traits,file="../data/lh.RData")
```

[Back to Top](#top)

## Author information

**Laurence Kell**.
[laurie\@seaplusplus.co.uk](mailto:laurie@seaplusplus.co.uk){.email}

## Acknowledgements

## Software Versions

**`r version$version.string`**

-   knitr: `r packageVersion('knitr')`

-   FLCore: `r packageVersion('FLCore')`

-   FLBRP: `r packageVersion('FLBRP')`

-   FLasher: `r packageVersion('FLasher')`

-   FLife: `r packageVersion('FLife')`

-   ggplotFL: `r packageVersion('ggplotFL')`

-   rfishbase: `r packageVersion('rfishbase')`

-   SPMpriors: `r packageVersion('SPMpriors')`

-   FishLife: `r packageVersion('FishLife')`

-   plyr: `r packageVersion('plyr')`

-   dplyr: `r packageVersion('dplyr')`

-   reshape: `r packageVersion('reshape')`

-   ggplot2: `r packageVersion('ggplot2')`

-   ggpubr: `r packageVersion('ggpubr')`

-   ggcorrplot: `r packageVersion('ggcorrplot')`

-   GGally: `r packageVersion('GGally')`

**Compiled**: `r date()`

[Back to Top](#top)

[Back to Top](#top)

# References {#references}

[Back to Top](#top)
