---
title: "**Annex: 01: IEO**"
subtitle: "Data Analysis of IEO Deepwater Shark Data"
author: "DELEGATION OF THE EUROPEAN UNION"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: SIOFA, deepwater sharks, hierchical framework
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
editor_options: 
  markdown: 
    wrap: 72
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir="html")})
---

```{r, knitr, echo=FALSE, warning=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               
               cache     =TRUE, 
               cache.path="cache/01_IEO/",
               
               fig.path  ="../tex/figs/01_IEO/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

options(digits=3)

iFig=0
iTab=0
```

# Introduction

This document provides a summary of the shark dataset, which includes information on shark captures such as species, length, weight, sex, maturity, and other relevant data. The aim is to understand the diversity of species, the types of data collected, and assess the quality and coverage of the dataset.

```{r setup}
# Start with a clean environment
rm(list = ls())

# Load libraries
library(openxlsx)  # For reading Excel files
library(dplyr)     # For data manipulation
library(tidyr)     # For data manipulation
library(ggplot2)   # For plotting
library(mapdata)   # For map-based data
library(viridis)   # For color scales
```

# Data Preparation

This section is dedicated to preparing the shark datasets for analysis. It includes loading the datasets, standardizing the variables, and ensuring the data is in a format suitable for analysis.

```{r data-preparation}
# Load data
load("../data/ieo.RData")

# Cpue dataset
cpue_ieo <- ieo$cpue %>% 
  filter(Arte == "LLS") %>% 
  filter(year > 2017) %>%
  rename(Gear = Arte, Year = year, Latitude = y3, Longitude = x3)

# Biological dataset
biol_ieo <- ieo$biol %>% 
  filter(year > 2017)

# Standardize sex variable for biological data:
table(biol_ieo$sex, useNA = "always")

# Assigning "2" to females
biol_ieo$sex[biol_ieo$sex == "2"] <- "F"

# Filtering out rows where 'sex' is "3", "4", "5", "H" or "N"
biol_ieo <- biol_ieo %>%
  filter(!sex %in% c("3", "4", "5", "H", "N"))

#Add categories for maturity

### Females
#1: immature; 2: maturing; 3: mature; 4: uterine developing; 5: differentiating; 6: expecting; 
#7: post-natal, spent; 8: Re-starting development

### Males
#1: immature; 2: developing; 3: mature; 4: active

table(biol_ieo$mat, useNA = "always")

# Convert 'mat' to numeric
biol_ieo$mat <- as.numeric(biol_ieo$mat)

# Define maturity descriptions for females
female_mat_desc <- c("immature", "maturing", "mature", "uterine developing", 
                     "differentiating", "expecting", "post-natal, spent", "re-starting development")

# Define maturity descriptions for males
male_mat_desc <- c("immature", "developing", "mature", "active")

# Update the biol_ieo dataset
biol_ieo <- biol_ieo %>%
  mutate(
    MaturityDescription = ifelse(sex == "F",
                                 ifelse(mat >= 1 & mat <= length(female_mat_desc), female_mat_desc[mat], NA),
                                 ifelse(sex == "M",
                                        ifelse(mat >= 1 & mat <= length(male_mat_desc), male_mat_desc[mat], NA),
                                        as.character(mat)))  # Default case
  )
```

# Maps of Catch Locations

This section focuses on visualizing and analyzing the geographical distribution of shark catches. Using maps, it aims to show where different shark species have been caught, the quantity of catches, and how these patterns vary based on the fishing gear used and over different years. This spatial analysis is crucial for understanding fishing hotspots, species distribution, and the impact of various fishing gears on different shark species.

## Catch weight per year and gear

We create map-based plots that combine geographic coordinates with catch data. These visualizations serve as a tool for understanding fishing patterns and their impact on shark populations, highlighting areas of intense fishing activity and shifts in fishing practices over time.

```{r location-maps, eval=FALSE}
# Summarize weight per year and gear
summarized_catch <- cpue_ieo %>%
  filter(!is.na(Latitude) & !is.na(Longitude) & !is.na(Year)) %>%
  group_by(species, Gear, Year, Longitude, Latitude) %>%
  summarize(TotalCatch = sum(Kilos, na.rm = TRUE))

# Add world map
world_map <- borders("worldHires", colour = "gray50", fill = "gray90")

# Variables for unique years, gears, and species
unique_years <- unique(summarized_catch$Year)
unique_gears <- unique(summarized_catch$Gear)
unique_species <- unique(summarized_catch$species)

# Settings for plot aesthetics
title_size <- 14  # Size for main title
axis_title_size <- 12  # Size for axis titles
legend_text_size <- 9  # Size for legend text

# Loop through each year and gear to create plots
for (year in unique_years) {
  for (gear in unique_gears) {
    # Filter data for the specific year and gear
    year_gear_data <- subset(summarized_catch, Year == year & Gear == gear)
    # Check if the year_gear_data is not empty
    if (nrow(year_gear_data) > 0) {
      # Create the plot with adjusted settings
      p <- ggplot() +
        world_map +
        geom_point(data = year_gear_data, aes(x = Longitude, y = Latitude, size = TotalCatch, color = species), alpha = 0.8) +
        coord_fixed(1.3, xlim = c(30, 100), ylim = c(-50, 5)) +
        scale_size_continuous(name = "Total Catch Weight") +
        scale_color_viridis(discrete = TRUE, name = "Species") +
        ggtitle(paste("Catch Weight in", year, "using", gear)) +
        theme_minimal() +
        theme(plot.background = element_rect(fill = "white", colour = "white"),
              panel.background = element_rect(fill = "white", colour = "white"),
              legend.position = "right",  # Position the legend at the bottom
              legend.background = element_rect(fill = "white", colour = "black"),
              legend.key = element_blank(),
              legend.text = element_text(size = legend_text_size),
              legend.title = element_text(size = legend_text_size),
              plot.title = element_text(size = title_size, face = "bold"),
              axis.title.x = element_text(size = axis_title_size, face = "bold"),
              axis.title.y = element_text(size = axis_title_size, face = "bold")) +
        scale_x_continuous(limits = c(30, 100)) +
        scale_y_continuous(limits = c(-50, 5)) +
        labs(title = paste("Catch Weight in", year, "using", gear), x = "Longitude", y = "Latitude")
      
      # Print the plot
      print(p)
    }
  }
}

```

```{r location-maps-1, fig.height=12,fig.width=10, eval=FALSE}

p=ggplot() +
        world_map +
        geom_point(data=summarized_catch, aes(x = Longitude, y = Latitude, size = TotalCatch, color = species), alpha = 0.8) +
        coord_fixed(1.3, xlim = c(30, 100), ylim = c(-50, 5)) +
        scale_size_continuous(name = "Total Catch Weight") +
        scale_color_viridis(discrete = TRUE, name = "Species") +
        #ggtitle(paste("Catch Weight in", year, "using", gear)) +
        theme_minimal() +
        theme(plot.background = element_rect(fill = "white", colour = "white"),
              panel.background = element_rect(fill = "white", colour = "white"),
              legend.position = "right",  # Position the legend at the bottom
              legend.background = element_rect(fill = "white", colour = "black"),
              legend.key = element_blank(),
              legend.text = element_text(size = legend_text_size),
              legend.title = element_text(size = legend_text_size),
              plot.title = element_text(size = title_size, face = "bold"),
              axis.title.x = element_text(size = axis_title_size, face = "bold"),
              axis.title.y = element_text(size = axis_title_size, face = "bold")) +
        scale_x_continuous(limits = c(30, 100)) +
        scale_y_continuous(limits = c(-50, 5)) +
        facet_wrap(~Year,ncol=2)
```

# Catch Composition

In this section, we analyse the shark catch data, focusing on how the catch varies both annually and across different geographic areas.

## Total catch by year

Here we examine the total shark catch over different years. By aggregating the total weight of catches annually, we aim to identify trends, such as increases or decreases in catch volumes over time.

```{r total-catch-year}

# Calculate total catch by year
total_catch_by_year <- cpue_ieo %>%
  group_by(Year) %>%
  summarize(TotalCatch = sum(Kilos, na.rm = TRUE))

# Bar chart of catch composition by year
ggplot(total_catch_by_year, aes(x = factor(Year), y = TotalCatch)) +
  geom_bar(stat = "identity", fill = "blue") +
  theme_minimal() +
  labs(x = "Year", y = "Total Catch (Kilos)", title = "Total Catch by Year")

```

## Catch composition by area

Here, we explore the catch composition across different fishing areas. By summarizing the total catch weight in each area, we gain an understanding of the geographical distribution of fishing efforts and shark abundance.

```{r catch-composition-area}
# Calculate total catch by area
catch_by_area <- cpue_ieo %>%
  group_by(Area) %>%
  summarize(TotalCatch = sum(Kilos, na.rm = TRUE)) %>%
  arrange(desc(TotalCatch))

# Bar chart of catch composition by area
ggplot(catch_by_area, aes(x = Area, y = TotalCatch, fill = Area)) +
  geom_bar(stat = "identity") +
  labs(title = "Catch Composition by Area", x = "Area", y = "Total Catch (Kilos)") +
  theme_minimal()

```

# Species Composition

This section is dedicated to analyzing the shark catch data from a species-centric viewpoint. It provides insights into the average catch weight of each species, the total catch composition broken down by species, and an exploration of species diversity across different fishing areas.

## Average catch weight by species

Here, we examine the average weight of catches for each shark species. This helps in identifying which species are predominantly heavier or lighter in the catches, potentially indicating their abundance and size in the fishing areas.

```{r average-catch-weight}

# Calculate average catch weight by species
avg_catch_weight_by_species <- cpue_ieo %>%
  group_by(species) %>%
  summarize(AverageCatchWeight = mean(Kilos, na.rm = TRUE))

# Create bar plot
ggplot(avg_catch_weight_by_species, aes(x = reorder(species, AverageCatchWeight), y = AverageCatchWeight)) +
  geom_bar(stat = "identity", fill = "cyan") +
  coord_flip() +  # Flipping the coordinates for easier reading
  theme_minimal() +
  labs(x = "Species (Scientific Name)", y = "Average Catch Weight", title = "Average Catch Weight by Species")

```

## Total catch composition by species

This analysis visualizes the total catch quantity for each shark species. It highlights which species are more frequently caught, thereby allowing us to assess the impact of fishing on different shark populations.

```{r species-catch-composition}

# Group weight for each species
catch_by_species <- cpue_ieo %>%
  group_by(species) %>%
  summarize(TotalCatch = sum(Kilos, na.rm = TRUE)) %>%
  arrange(desc(TotalCatch))

# Bar chart of catch composition by species
ggplot(catch_by_species, aes(x = reorder(species, TotalCatch), y = TotalCatch)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(title = "Catch Composition by Species", x = "Species", y = "Total Catch (Kilos)") +
  theme_minimal()

```

## Species composition by area

In this subsection, we investigate the presence and prevalence of shark species in various fishing areas. This spatial distribution analysis helps in understanding the habitat preferences of different shark species.

```{r species-composition-area}

# Count of species in each area
species_count_by_area <- cpue_ieo %>%
  group_by(Area, species) %>%
  summarize(Count = n(), .groups = 'drop')

# Grouped bar chart
ggplot(species_count_by_area, aes(x = Area, y = Count, fill = species)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Species Presence in Each Area", x = "Area", y = "Count of Occurrences") +
  theme_minimal() +
  scale_fill_viridis(discrete = TRUE)

```

# Temporal Analysis

This section focus on the temporal dynamics of shark catches, providing insights into catch trends over different time scales. Understanding these trends is crucial for assessing the health and sustainability of shark populations over time.

## Overall catch trends on a yearly basis

This analysis focuses on how the total shark catch has varied year by year. The line graph highlights any significant increases or decreases in catch over the years, which can be indicative of changes in shark populations, fishing effort, or regulations.

```{r yearly-catch-trend}

# Yearly catch trend
yearly_trend <- cpue_ieo %>%
  group_by(Year) %>%
  summarize(TotalCatch = sum(Kilos, na.rm = TRUE))

# Line graph for yearly trends
ggplot(yearly_trend, aes(x = Year, y = TotalCatch)) +
  geom_line(color = "blue") +
  geom_point(color = "red") +
  labs(title = "Yearly Catch Trend", x = "Year", y = "Total Catch (Kilos)") +
  theme_minimal()

```

## Overall catch trends on a montly basis

Here, we extend the temporal analysis to a finer scale by examining monthly trends in shark catches. This can reveal seasonal patterns in fishing activities or shark abundance, offering a more detailed understanding of the dynamics within each year.

```{r monthly-catch-trend, echo=TRUE, message=FALSE, warning=FALSE}

# Montly catch trend
monthly_trend <- cpue_ieo %>%
  group_by(Year, month) %>%
  summarize(TotalCatch = sum(Kilos, na.rm = TRUE))

# Line graph for monthly trends
ggplot(monthly_trend, aes(x = month, y = TotalCatch, group = Year, color = factor(Year))) +
  geom_line() +
  geom_point() +
  labs(title = "Monthly Catch Trend", x = "Month", y = "Total Catch (Kilos)", color = "Year") +
  theme_minimal()

```

## Species temporal trend analysis

In this part, we look at the yearly catch trends by species. This species-specific temporal analysis helps in identifying which shark species have seen significant changes in their catch over time, potentially signaling shifts in species-specific abundance or susceptibility to fishing pressures.

```{r species-temporal-trend, echo=TRUE, message=FALSE, warning=FALSE}

# Aggregate by year and species
species_yearly_trend <- cpue_ieo %>%
  group_by(Year, species) %>%
  summarize(TotalCatch = sum(Kilos, na.rm = TRUE))

# Line graph for species-specific yearly trends
ggplot(species_yearly_trend, aes(x = Year, y = TotalCatch, color = species)) +
  geom_line() +
  labs(title = "Yearly Catch Trend for Each Species", x = "Year", y = "Total Catch (Kilos)", color = "Species") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

# Species Distribution By Length

This section provides an in-depth look at the distribution of shark sizes within the dataset. Analyzing the length of sharks is crucial for understanding their growth patterns, population structure, and the impact of fishing on different size classes.

## Overall sizes distribution

This histogram offers a broad overview of the sizes of sharks across all species in the dataset. It reveals the most common size ranges and any notable size-related trends or anomalies.

```{r sizes-distribution}

# Bar plot on sizes distributions
ggplot(subset(biol_ieo, !is.na(len)), aes(x = len)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  #facet_wrap(~Scientific_name) +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white", colour = "white"),
        panel.background = element_rect(fill = "white", colour = "white")) +
  labs(title = "Distribution of Fish Sizes", x = "Length", y = "Count")

```

## Average length by species

The bar chart here displays the average length of each shark species, allowing for a comparative analysis of size across different species. This can be particularly informative for understanding species-specific growth patterns and identifying outliers.

```{r Average-length-species}

# Calculate average length for each species
avg_length_by_species <- biol_ieo %>%
  group_by(species) %>%
  summarize(AverageLength = mean(len, na.rm = TRUE))

# Plot the average lenghts 
ggplot(avg_length_by_species, aes(x = reorder(species, AverageLength), y = AverageLength)) +
  geom_bar(stat = "identity", fill = "blue") +
  coord_flip() +  # For better readability of species names
  theme_minimal() +
  labs(title = "Average Length by Species", x = "Species", y = "Average Length (cm)")

```

## Distribution of fish sizes by species

This detailed histogram, with each facet representing a different species, provides a species-specific view of size distribution. It helps in identifying how size distribution varies across species, which is important for species-specific management and conservation strategies.

```{r Fish-Sizes-Species}

# Bar plot on the fish sizes distributions for each species 
ggplot(subset(biol_ieo, !is.na(len)), aes(x = len)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  facet_wrap(~species, scales = "free_x") +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white", colour = "white"),
        panel.background = element_rect(fill = "white", colour = "white")) +
  labs(title = "Distribution of Fish Sizes by Species", x = "Length", y = "Count")

```

## Size distribution by sex

The histogram here breaks down the size distribution based on sex, offering insights into potential sexual dimorphism in size. This can be crucial for understanding reproductive strategies and population dynamics.

```{r size-distribution-sex}

# Histogram for size distribution by sex
ggplot(subset(biol_ieo, !is.na(sex) & !is.na(len)), aes(x = len, fill = sex)) +
  geom_histogram(binwidth = 5, position = "dodge") +
  labs(title = "Size Distribution by Sex", x = "Length", y = "Count", fill = "Sex") +
  theme_minimal()

```

## Size distribution by maturity stage

Finally, the box plot in this section looks at the size distribution across different maturity stages. This visualization helps in understanding the growth trajectory of sharks and how their size correlates with their maturity stage, which is essential for effective population management. The plot is further faceted by sex to provide a more nuanced view of these dynamics.

```{r size-distribution-maturity-stage}

# Box plot for size distribution by maturity stage
ggplot(subset(biol_ieo, !is.na(MaturityDescription) & !is.na(len)), aes(x = MaturityDescription, y = len, fill = MaturityDescription)) +
  geom_boxplot() +
  labs(title = "Size Distribution by Maturity Stage", x = "Maturity Stage", y = "Length", fill = "Maturity Stage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotates x-axis labels
  theme(legend.position = "none") +
  facet_wrap(~sex)

```

# Gender Distribution

This section focus on the gender distribution within the shark dataset, providing insights into the male-to-female ratios among different shark species. Understanding gender distribution is crucial for assessing reproductive capacities and potential vulnerabilities of shark populations.

## Overall gender distribution

The bar plot in this subsection gives an overview of the gender distribution across all shark species. This helps in understanding the overall balance between male and female sharks in the dataset, which is vital for population studies and conservation efforts.

```{r gender-distribution}

# Aggregating data by species and gender
gender_distribution <- biol_ieo %>%
  filter(!is.na(sex)) %>%
  group_by(species, sex) %>%
  summarize(Count = n(), .groups = 'drop')

# Creating a faceted bar plot
ggplot(gender_distribution, aes(x = sex, y = Count, fill = sex)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white", colour = "white"),
        panel.background = element_rect(fill = "white", colour = "white")) +
  labs(title = "Gender Distribution", x = "Gender", y = "Count") +
  scale_fill_brewer(palette = "Spectral", name = "Gender")

```

# Gender ratio by species

This plot takes a species-specific approach to analyzing gender distribution. By breaking down the male and female counts for each species, it highlights any disparities in gender ratios among different shark species. Such information is essential for species-specific management practices, especially for species that might have skewed gender ratios due to fishing pressures or other environmental factors.

```{r species-gender-ratio}

# Gender ratio by species
gender_ratio_by_species <- biol_ieo %>%
  group_by(species) %>%
  summarize(FemaleCount = sum(sex == "F", na.rm = TRUE),
            MaleCount = sum(sex == "M", na.rm = TRUE))

gender_ratio_by_species <- gender_ratio_by_species %>%
  pivot_longer(cols = c("FemaleCount", "MaleCount"), names_to = "Gender", values_to = "Count")

# Bar plot on gender ratio by species
ggplot(gender_ratio_by_species, aes(x = species, y = Count, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Gender Ratio by Species", x = "Species", y = "Count") +
  scale_fill_manual(values = c("FemaleCount" = "pink", "MaleCount" = "blue"))

```

# Maturity

This section focuses on the maturity stages of sharks, a critical aspect for understanding their reproductive status and health. The analysis is conducted both at the species level and by sex, providing a comprehensive view of the maturity distribution.

## Proportion of each maturity stage within each species

In this subsection, a stacked bar plot is used to depict the distribution of various maturity stages across different shark species. This visualization allows for a quick assessment of the predominant maturity stages in each species, which is crucial for understanding their reproductive cycles and potential for recruitment.

```{r proportion-maturity-stage}

# Aggregate species by maturity stage and sex
maturity_count_by_species <- biol_ieo %>%
  filter(!is.na(MaturityDescription), !is.na(sex)) %>%
  group_by(species, MaturityDescription, sex) %>%
  summarize(Count = n(), .groups = 'drop')

# Bar plot showing the proportion of each maturity stage within each species.
ggplot(maturity_count_by_species, aes(x = species, y = Count, fill = MaturityDescription)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(title = "Maturity Stage Counts by Species", x = "Species", y = "Count") +
  scale_fill_brewer(palette = "Paired") +
  coord_flip()  # Optional for better readability of species names


```

## Proportion of each maturity stage within each species, separated by sex

This plot separates the data by sex, providing a gender-specific view of maturity stages for each species. This approach is particularly useful for identifying any sex-based differences in maturity stages. The facet wrap creates individual panels for each sex, making it easier to compare and contrast these patterns across species.

```{r proportion-maturity-sex}

# Bar plot showing the proportion of each maturity stage within each species, separated by sex.
ggplot(maturity_count_by_species, aes(x = species, y = Count, fill = MaturityDescription)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~ sex, scales = "free_y") +  # Separate panels for each sex
  theme_minimal() +
  labs(title = "Maturity Stage Counts by Species and Sex", x = "Species", y = "Count") +
  scale_fill_brewer(palette = "Paired") +
  coord_flip()  # Optional for better readability of species names

```
