---
title: "**Annex: 02:**"
subtitle: "Life history priors"
author: "DELEGATION OF THE EUROPEAN UNION"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: SIOFA, deepwater sharks, hierchical framework
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
editor_options: 
  markdown: 
    wrap: 72
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir="html")})
---


```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               
               cache     =TRUE,
               cache.path="cache/02/",
               
               fig.path  ="../figs/02/",
               fig.width =10,
               fig.height=8,
               dev       ="png")

iFig=0
iTab=0
```

```{r, eval=FALSE}
install.packages("https://cran.r-project.org/src/contrib/Archive/rfishbase/rfishbase_3.0.0.1.tar.gz", repos=NULL, type='source')

R CMD INSTALL "--no-staged-install" rfishbase
```


```{r, pkgs}
library(ggplot2)
library(ggpubr)
library(GGally)
library(ggcorrplot)
library(ggpubr)

library(statcomp)
library(scales)

library(plyr)
library(dplyr)
library(reshape)

library(FLCore)
library(FLBRP)
library(FLasher)
library(FLife)
library(ggplotFL)

library(rfishbase)
library(SPMpriors)
library(FishLife)

library(R.devices)
```

```{r, source}
source("../source/themes.R") 
source("../source/fl_mcupd_fn.R") 
```


# Data
```{r, spp-list}
load("../data/spp.RData")
load("../data/lh.RData")

load("../data/ieo.RData")
load("../data/siofa.RData") 
```


```{r, cor, fig.height=8, fig.width=8}
lh=cast(priors,Genus+Species~trait,value="data")  
lh=transmute(lh,spp=paste(Genus,Species),linf=exp(Loo),k=exp(K),l50linf=exp(Loo-Lm),l50=exp(Lm),G=G,M=exp(M),r=r)

ggpairs(lh[,-(1)],
              #mapping = ggplot2::aes(color=rank(dat$hmsy)<6))+
     lower = list(continuous = wrap(my_smooth)), 
     diag=list(continuous=wrap(my_density,alpha=0.2)))+
     theme_bw(16)+theme(legend.position = "none",  
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA", angle=90), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"),
                    
                    axis.title.x = element_blank(), 
                    axis.text.x  = element_blank(), 
                    axis.ticks.x = element_blank(), 
                    
                    plot.margin = unit(c(0, 0, 0, 1), "lines"),
                    panel.background = element_rect(fill   ="NA", colour="NA"), 
                    #panel.border     = element_rect(fill   ="NA", colour="NA"), 
                    panel.grid.major = element_line(colour ="NA"), 
                    panel.grid.minor = element_line(colour ="NA"))
```

**Figure `r iFig=iFig+1; iFig`** Relationships between life history relationships, individual growth parameters $L_{\infty}$ $k$, length at maturity ($L_{50}$), and their ratio ($L{50}/L_{\infty}$) and derived quantities natural mortality ($M$), population grpwth rate ($r$) and generation time ($G$).


```{r, priors, fig.height=8,fig.width=8}
m=ddply(subset(priors,trait=="M"), .(Genus,Species,trait), with,
           data.frame(iter=seq(100),dev=exp(rnorm(100,data,var)),mean=data,sd=var^0.5))
r=ddply(subset(priors,trait=="r"), .(Genus,Species,trait), with,
           data.frame(iter=seq(100),dev=rnorm(100,data,var),mean=data,sd=var^0.5))
rfpts=rbind(cbind(Derivation="M",mutate(m,Fmsy=0.41*dev,Flim=1.5*Fmsy,Fcrash=2*Fmsy)),
            cbind(Derivation="r",mutate(r,Fmsy=0.5*dev, Flim=1.5*Fmsy,Fcrash=2*Fmsy))) 
gt=subset(priors,trait=="G"&Species%in%unique(r$Species)&Genus%in%unique(r$Genus))
```

```{r, m, fig.height=8,fig.width=8}
ggdensity(m,x="dev",fill="Species",alpha=0.25)+
  facet_wrap(Genus~.,scales="free_y")+
  scale_x_continuous(limits=c(0,.5),breaks=c(0,0.25,0.5))+    
  theme(legend.position = "none", 
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA"), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"))
```

**Figure `r iFig=iFig+1; iFig`** Estimates of Natural Mortality (M) based on life-history theory

```{r, r, fig.height=8,fig.width=8}
ggdensity(r,x="dev",fill="Species",alpha=0.25)+
  facet_wrap(Genus~.,scales="free_y")+
  scale_x_continuous(limits=c(0,.5),breaks=c(0,0.25,0.5))+    
  theme(legend.position = "none", 
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA"), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"))
```

**Figure `r iFig=iFig+1; iFig`** Estimates of population growth $r$ based on life-history theory


```{r, fmsy, fig.height=12,fig.width=10}
ggdensity(rfpts,x="Fmsy",fill="Derivation")+
  facet_wrap(Genus~.,scale="free_y")+
  scale_x_continuous(limits=c(0,0.3),breaks=c(0.1, 0.2, 0.3))+
  xlab(expression(F[MSY]))+
    theme(legend.position = "none", 
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA"), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"))
```

**Figure `r iFig=iFig+1; iFig`** Estimates of $F_{MSY}$ based on life-history theory



```{r, box-m, fig.height=12,fig.width=10}
h<-function(x) 1-exp(-x)

fn<-function(x) {
  r=quantile(x, probs=c(0.05, 0.33, 0.5, 0.66, 0.66))
  names(r)=c("ymin", "lower", "middle", "upper", "ymax")
  r}

spp=inventory[!duplicated(inventory$spp),"spp"]

dat=subset(rbind(cbind("Assumption"="M",transform(m,x=h(dev*0.41))),
                 cbind("Assumption"="r",transform(r,x=h(dev*0.5)))),
           paste(Genus,Species)%in%spp)
ggplot(dat,aes(x,y=paste(Genus,Species)))+  
   stat_summary(fun.data=fn, geom="boxplot")+  
   labs(title = "Estimates of Fmsy (% per Year)")+
   facet_grid(Genus~Assumption,scale="free",space="free")+
   theme_bw(16)+ 
   theme(legend.position="right",
        #axis.title.x = element_text(colour='NA'), 
        axis.ticks =   element_line(colour="NA"),
        axis.text.x=element_text(angle=-45, hjust=0),
        strip.text.y = element_blank())+
   xlab("Natural mortality rate by species obtained from life history theory")+ylab("")+
   scale_x_continuous(breaks=c(0,0.05,0.10,0.15),labels=percent)
```

**Figure `r iFig=iFig+1; iFig`** Estimates of $F_{MSY}$ based on life-history theory, boxes indicate the 33rd to 66th percentiles and the median, the lower whisker shows the 5th percentile.



```{r, box-doubling, fig.height=12,fig.width=10}
fn<-function(x) {
  r=quantile(x, probs=c(0.05, 0.33, 0.5, 0.66, 0.95))
  names(r)=c("ymin", "lower", "middle", "upper", "ymax")
  r}

#gt=subset(mrag,trait=="ln_G")

lvl=priors[!duplicated(priors[,c("Genus","Species")]),c("Genus","Species","level")]
r  =merge(r,lvl)

ggplot(r,aes(x=log(2)/dev*2,y=paste(Genus,Species),fill=ac(level)))+  
   stat_summary(fun.data=fn, geom="boxplot")+  
   geom_point(aes(x=data,y=paste(Genus,Species)),data=gt,col="red")+
   labs(title = "Population Doubling Time at Bmsy")+
   facet_grid(Genus~.,scale="free",space="free")+
   theme_bw(16)+ 
   theme(legend.position="right",
        #axis.title.x = element_text(colour='NA'), 
        axis.ticks =   element_line(colour="NA"),
        axis.text.x=element_text(angle=-45, hjust=0),
        strip.text.y = element_blank())+
   geom_hline(aes(yintercept=0))+
   theme(axis.text.x=element_text(angle=-45, hjust=0))+
   xlab("Population Doubling Time")+ylab("")+
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA))
```

**Figure `r iFig=iFig+1; iFig`** Estimates of $F_{MSY}$ based on life-history theory, boxes indicate the 33rd to 66th percentiles and the median, the lower whisker shows the 5th percentile.


```{r, eval=FALSE}
library(ggplot2)  
library(RColorBrewer)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(lubridate)
library(plyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(gridExtra)

library(aquamapsdata)
library(raquamaps)
library(gfwr)

library(leaflet)
library(raster)
library(aquamapsdata)

library(aquamapsdata)
#download_db(force = TRUE)
default_db("sqlite")

ss =read_sf(file.path("/home/laurie/pCloudDrive/shapeFiles/SS_Boundary/Sargasso_Sea_Geographical_Area_of_Collaboration.shp"))
ssE=read.csv("/home/laurie/Dropbox/Sargasso FFEM Shared 2021/06 Deliverables/analysis/inputs/gfw/bq-results-20221012-060605-1665555242269.csv")

world <- ne_countries(scale = "medium", returnclass = "sf")

#d2p <- nd %>% 
#  group_by(lat, lon, gear) %>% 
#  summarise(fishing_time = sum(fishing_time))

ss =read_sf(file.path("/home/laurie/pCloudDrive/shapeFiles/SS_Boundary/Sargasso_Sea_Geographical_Area_of_Collaboration.shp"))
ssE=read.csv("/home/laurie/Dropbox/Sargasso FFEM Shared 2021/06 Deliverables/analysis/inputs/gfw/bq-results-20221012-060605-1665555242269.csv")

world <- ne_countries(scale = "medium", returnclass = "sf")

iccat=read_sf("/home/laurie/pCloudDrive/shapeFiles/ICCAT/ICCAT_draft_ecoregions_May 2022.shp")

ggplot(subset(ssE,geartype=="trawlers")) +
    geom_tile(aes(x=lon_bin_center, y=lat_bin_center, fill=log10(fishing_hours))) +
    scale_fill_distiller(palette = "Blues", direction = 1) +
    geom_sf(data = ss, fill = NA)+
    #geom_s f(data = iccat, fill = NA, col="red")+
    geom_sf(data = world) +
  theme_bw()+
  labs(x = "Longitude", y= "Latitude") +
      coord_sf(expand = FALSE,
             xlim=c(c(-40,-80)), ## set the plot longitude extents
             ylim=c(20,40)) ## set the plot latitude extents

key <- am_search_fuzzy("Illex illecebrosus")$key
ras <- am_raster(key)

# show the native habitat map
am_map_leaflet(ras, title = "Shortfin Squid") %>%
leaflet::fitBounds(lng1 =-75, lat1=10, lng2=-30, lat2=55) %>%
  addPolygons(data=ss, weight = 2)
```

