---
title: "**Annex: 12**"
subtitle: "Maturity Ogive for Centroscymnus coelolepis"
author: "DELEGATION OF THE EUROPEAN UNION"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: SIOFA, deepwater sharks, hierchical framework
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
editor_options: 
  markdown: 
    wrap: 72
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir="html")})
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo=FALSE)

library(knitr)
opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path="cache/12/",
               fig.path  ="../figs/12",
               fig.width =6,
               fig.height=6,
               dev       ="png")
iFig=0
iTab=0
```


```{r, libs, cache=FALSE}
library(openxlsx)  # For reading Excel files
library(lubridate) # For handling date and time
library(plyr)      # For data manipulation
library(dplyr)     # For data manipulation
library(reshape)   # For data manipulation
library(ggplot2)   # For plotting
library(ggpubr)    # For plotting
library(mapdata)   # For map-based data
library(viridis)   # For color scales, better for visibility

library(xtable)

library(FLCore)
library(ggplotFL)
```


```{r, theme, cache=FALSE}
theme_my<-function(size=16) {
  theme_minimal() +
    theme(text=element_text(size=size),
          plot.title=element_text(face="bold", size=size+2),
          axis.title=element_text(size=size-1),
          legend.position="bottom",
          legend.title=element_blank(),
          panel.grid.major=element_line(color="gray80"),
          panel.grid.minor=element_blank())}

theme_set(theme_my())  
```

```{r, setwd, eval=FALSE}
setwd("~/Desktop/active/siofa-haf/Rmd")
setwd("C:/Users/lauri/OneDrive/Desktop/active/siofa-haf/Rmd")
```


```{r, mat2}
pDogMorph=read.xlsx("../data/inputs/maturity_data.xlsx", "morphometric", 
                    startRow=1, detectDates=TRUE, skipEmptyRows=TRUE)
pDogGonad=read.xlsx("../data/inputs/maturity_data.xlsx", "gonad", 
                    startRow=1, detectDates=TRUE, skipEmptyRows=TRUE)
mat=rbind.fill(cbind(Type="Morph",pDogMorph),
               cbind(Type="Gonad",pDogGonad))
mat$sex =factor(mat$sex,labels=c("Female","Male"),levels=c("F","M"))
mat$Type=factor(mat$Type)
```

```{r, mat2-ieo}
ggplot(mat)+
  geom_line(aes(len,fitted,col=sex,linetype=Type))+
  xlab("Length (cm)")+ylab("Proportion Mature")
```

**Figure `r iFig=iFig+1; iFig`** .


#### Fit logistic regression model
```{r, mat2-fit}
model=glm(mature~len+sex+len:sex+len:Type, data=mat, family=binomial)
mat$hat=predict(model,type="response")

ggplot(mat, aes(x=len, y=hat, color=sex, linetype=Type)) + 
  geom_line() + 
  labs(x="Length", y="Predicted Maturity Probability") + 
  scale_y_continuous(limits=c(0, 1)) + 
  theme_minimal()
```


**Figure `r iFig=iFig+1; iFig`** .


### 1. Summary of Model Fit
- Use `summary(model)` to get an overview of the model fit, including coefficients, standard errors, z-values, and P-values for each predictor.

```{r}
summary(model)  
```

### 2. Residuals and Model Fit
- **Residuals Plot**: Plot residuals to look for patterns. For logistic regression, deviance residuals can be informative.

```{r}
mat$rsdl=residuals(model, type="deviance")

ggplot(mat)+
  geom_point(aes(len,rsdl))+
  facet_grid(sex~Type)
```


**Figure `r iFig=iFig+1; iFig`** .


- **Hosmer-Lemeshow Test**: Test for goodness of fit specifically designed for logistic regression.
```{r}
library(ResourceSelection)
hoslem.test(model$y, fitted(model))
```

### 3. Influence Measures
- **Cook's Distance**: Identify influential observations based on Cook's distance.

### 3. Influence Measures
- **Cook's Distance**: Identify influential observations based on Cook's distance.
```{r, eval=TRUE, echo=FALSE}
mat$cdist=cooks.distance(model)

ggplot(mat) +
  geom_linerange(aes(len,ymin=0,ymax=cdist))+
  labs(title="Influence Plot Using Cook's Distance",
       x="Length (cm)",
       y="Cook's Distance") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1))+
  facet_grid(sex~Type)
```


**Figure `r iFig=iFig+1; iFig`** .


- **Leverage and Influence Plots**: Identify observations with high leverage or influence on the model estimation.
```{r}
library(car)
influencePlot(model, id.method="identify", main="Influence Plot", sub="Circle size is proportional to Cook's Distance")
```

To create an influence plot using `ggplot2` that mimics the functionality of the `car::influencePlot`, you'll need to manually calculate the elements you want to visualize: leverage, standardized residuals, and Cook's distance. Here's how you can do it with a logistic regression model as an example. This approach involves extracting the necessary statistics from the model and then plotting them using `ggplot2`.

First, ensure you have your logistic regression model fitted. For this example, let's assume your model is stored in a variable named `model`. You'll need the `broom` and `ggplot2` packages as well.

```{r}
library(broom)

# Calculate leverage (hat values)
model_leverage <- hatvalues(model)

# Calculate standardized residuals
model_std_resid <- rstandard(model)

# Calculate Cook's distance
model_cooks_d <- cooks.distance(model)

# Create a dataframe for plotting
influence_data <- data.frame(
  Leverage=model_leverage,
  StdResiduals=model_std_resid,
  CooksDistance=model_cooks_d
)

# Generate the plot
ggplot(cbind(mat,influence_data), aes(x=Leverage, y=StdResiduals, size=CooksDistance)) +
  geom_point(alpha=0.6) +
  scale_size_continuous(range=c(1, 12), guide='none') + # Adjust size scale and hide legend
  labs(title="Influence Plot",
       subtitle="Circle size is proportional to Cook's Distance",
       x="Leverage",
       y="Standardized Residuals") +
  theme_minimal()+
  facet_grid(sex~Type)
```

**Figure `r iFig=iFig+1; iFig`** .

This script does the following:
  - Calculates the leverage, standardized residuals, and Cook's distance for each observation in your logistic regression model.
- Constructs a dataframe `influence_data` that contains these values.
- Uses `ggplot2` to create a scatter plot where the x-axis is leverage, the y-axis is standardized residuals, and the size of each point represents Cook's distance.

Adjust the `scale_size_continuous` function as needed to better fit your data. The `guide='none'` argument hides the size legend since it generally clutters the plot. This approach gives you a detailed and customizable influence plot using `ggplot2`.

Creating an influence plot with `ggplot2` requires calculating influence measures like Cook's distance, leveraging the model you've fitted, and then plotting these measures. The influence plot visually identifies points that have a significant impact on the model's estimates. Here's how you can create an influence plot for your logistic regression model using `ggplot2` in R:
  
  1. **Calculate Cook's Distance**: Cook's distance is a measure used to estimate the influence of each data point. In logistic regression, it identifies points that, if removed, would change the model's parameters significantly.

2. **Prepare the Data**: Combine the Cook's distance with your original dataset or create a new dataframe with the Cook's distance and leverage values for each observation.

3. **Plot Using `ggplot2`**: Create the plot using `ggplot2`, highlighting observations with high influence.

Here's an example code snippet that demonstrates this process:
  
  ### 4. Multicollinearity
  - **Variance Inflation Factor (VIF)**: Check for multicollinearity among predictors.
```{r}
vif(model)
```

### 5. Model Assumptions and Fit
- **Check Linearity**: The logit link assumes a linear relationship between log odds and predictors. Use plots or Generalized Additive Models (GAMs) to assess this.
```{r}
library(mgcv)
gam_model=gam(mature ~ s(len) + sex + Type, family=binomial, data=mat)
plot(gam_model)
```

```{r}
# Load necessary libraries
library(mgcv)
library(ggplot2)

# Assuming your GAM model is named gam_model
gam_model <- gam(mature ~ s(len) + sex + Type, family=binomial(), data=mat)

# Extracting the smooth term for 'len'
smooth_len <- plot(gam_model, select=1, se=TRUE)[[1]] # 'select=1' selects the first smooth term, usually it's s(len)

# Convert to a data frame for ggplot
df_smooth_len <- data.frame(
  len   =smooth_len$x,
  fitted=smooth_len$fit,
  lower =smooth_len$fit - 2 * smooth_len$se,
  upper =smooth_len$fit + 2 * smooth_len$se
)

# Plot using ggplot2
ggplot(df_smooth_len, aes(x=len)) +
  geom_ribbon(aes(ymin=lower, ymax=upper), fill="lightblue", alpha=0.5) +
  geom_line(aes(y=fitted), color="blue") +
  labs(title="Effect of Length on Maturity", x="Length", y="Estimated Effect") +
  theme_minimal()

```


**Figure `r iFig=iFig+1; iFig`** .


- **Check for Overdispersion**: This is less of a concern in binary logistic regression but can be relevant for count data.

### 6. Predicted vs. Observed
- **Predicted vs. Observed**: Compare the predicted probabilities to the observed outcomes to evaluate the model's performance.

```{r,eval=FALSE}
fitted_probs=predict(model, type="response")
plot(pDogGonad$mature, fitted_probs, xlab="Observed", ylab="Predicted", main="Predicted vs Observed")
abline(0,1)
```


**Figure `r iFig=iFig+1; iFig`** .
