---
title: "Stock Assessment Framework for Deepwater Sharks in SIOFA"
author: "DELEGATION OF THE EUROPEAN UNION"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
#  powerpoint_presentation:
  html_document:
    toc: true
    toc_float: true
    #toc_depth: 2
    collapsed: true
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, lib}
library(openxlsx)  # For reading Excel files
library(lubridate) # For handling date and time
library(plyr)      # For data manipulation
library(dplyr)     # For data manipulation
library(reshape)   # For data manipulation
library(ggplot2)   # For plotting
library(mapdata)   # For map-based data
library(viridis)   # For color scales, better for visibility
```

```{r, eval=FALSE}
setwd("/home/laurie/Desktop/projects/siofa/Rmd")
```

```{r, data}
# Data Preparation
# Load data from an Excel file into data frames
cpue = read.xlsx("../data/inputs/SIOFA_DWS-2023-01-Data.xlsx", "dws_data_catch", 
                 startRow = 1, detectDates = TRUE, skipEmptyRows = TRUE)
biol = read.xlsx("../data/inputs/SIOFA_DWS-2023-01-Data.xlsx", "dws_data_bio_sampling",
                 startRow = 1, detectDates = TRUE, skipEmptyRows = TRUE)

# Adjusting time format
adjust_time_format<-function(time)
  format(as.POSIXct(time * 86400, origin = "1970-01-01", tz = "UTC"), "%H:%M:%S")

cpue$SettingTime=adjust_time_format(cpue$SettingTime)
cpue$HaulingTime=adjust_time_format(cpue$HaulingTime)
cpue$Time       =adjust_time_format(cpue$Time)
cpue$Year       =year(cpue$Date)

# Data Visualization and Analysis Functions

catchSmry<-function(data) {
  # Filter out NAs
  filtered=filter(data, !is.na(Latitude), !is.na(Longitude), !is.na(Year))
  # Group data
  group_by(filtered, speciesScientificName, Gear, Year, Latitude, Longitude)}

cSmry=catchSmry(cpue)
```

## Introduction

-   Overview of deepwater shark species within SIOFA's jurisdiction
-   Highlighting their vulnerability and importance of sustainable management

## Challenges

-   Discussing the challenges in assessing deepwater shark populations
-   Data limitations and life-history characteristics

## Framework Overview

-   Introducing the hierarchical, risk-based assessment framework proposed

## Data Sources

-   Detailing novel data collection methods like tagging and eDNA

## Adaptive Management Strategies

-   Need for flexible management approaches that evolve with new data

## Model Validation

-   Importance of using validated models for stock assessment

## Stakeholder Engagement

-   Role of engaging stakeholders in the management process

## Management Measures

-   Potential measures including spatial management and gear modification

## Conclusion

-   Reiterating the importance of the proposed framework
