---
title: "**SIOFA Sharks**"
subtitle: "Annex 5:  Catch and Effort Analysis for Centroscymnus coelolepis"
author: "DELEGATION OF THE EUROPEAN UNION"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
mathjax: TRUE
fig_width: 6
fig_height: 4
tags: FLR FLCore SIOFA
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
#bibliography: /home/laurie/pCloudDrive/refs.bib
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path="cache/cpue/",
               fig.path  ="../figs/cpue/",
               fig.width =12,
               fig.height=8,
               dev       ="png")
iFig=0
```


```{r setup}
library(openxlsx)
library(lubridate)
library(dplyr)
library(ggplot2)
library(mapdata)
library(viridis)
```

```{r setwd, eval=FALSE}
setwd("M:/My Documents/WMR-2024/Deep-water sharks SIOFA/Data/")
setwd("../Desktop/active/siofa-haf/Rmd")
```

## SIOFA dataset
The dataset contains information on shark catches. In this document we focus on the Portugese dogfish (Centroscymnus coelolepsis). Catch-per-unit-of-effort (CPUEs) are calculated for each gear type using different measures of fishing effort depending on the gear. For the different types of trawlers, CPUE is calculated using the tow duration (hours). For gillnets and entangling nets, the net length (m) is used as soaktime was not available for this gear type. For longlines and vertical lines, the number of hooks set was used and, lastly, for set longlines we used soaktime (hours). The time series is from 2003 - 2022, with low data available in 2003 and 2005, and no data available for 2006 and 2007.

```{r data}
# Data preparation
# Load data from excel file into data frames
df1  <- read.xlsx("../data/inputs/SIOFA_DWS-2023-01-Data.xlsx", sheet=2, 
                  startRow=1,detectDates=TRUE, skipEmptyRows=TRUE) %>% 
          filter(speciesScientificName %in% "Centroscymnus coelolepis") #/ Filter species

# Adjust the time format
df1$SettingTime <- as.POSIXct(df1$SettingTime * 86400, origin = "1970-01-01", tz = "UTC")
df1$SettingTime <- format(df1$SettingTime, "%H:%M:%S")
df1$HaulingTime <- as.POSIXct(df1$HaulingTime * 86400, origin = "1970-01-01", tz = "UTC")
df1$HaulingTime <- format(df1$HaulingTime, "%H:%M:%S")
df1$Time <- as.POSIXct(df1$Time * 86400, origin = "1970-01-01", tz = "UTC")
df1$Time <- format(df1$Time, "%H:%M:%S")

# Extract year from date
df1$Year <- year(df1$Date)
```


```{r CPUEs}
# Calculate CPUE for the different gear types
df <- df1 %>% 
       mutate(CPUE = ifelse(Gear %in% c("Bottom trawls (nei)","Midwater trawls (nei)","Single boat bottom otter trawls","Trawls (nei)"),
                            catchWeight/(TowDuration_minutes/60),
                     ifelse(Gear %in% c("Gillnets and entangling nets (nei)"),
                            catchWeight/NetLength_m,
                     ifelse(Gear %in% c("Longlines (nei)","Vertical lines"),
                            catchWeight/(Hooks_set/1000),
                     ifelse(Gear %in% c("Set longlines"),
                            catchWeight/(SoakTime_minutes/60),NA)))))

# Remove CPUE = NA
df <- df %>% filter(!is.na(CPUE))

# Remove negative soak time (Didn't work with filter)
idx <- which(df$SoakTime_minutes < 0)
df <- df[-idx,]

# Remove missing lon/lat -> this is 1 observation
df <- df %>% filter(!is.na(Longitude))

# Overview of total CPUE (sum) per year/gear
df %>% group_by(Year,Gear) %>% summarise(sum_cpue = sum(CPUE)) %>% tidyr::spread(Gear,sum_cpue)

```

The following maps show the spatio-temporal distribution of CPUE for Portugese dogfish in the SIOFA waters for each of the observed records. The species is found in higher densities on the latitudinal degrees between -30 and -35. Though these maps give some indication on the abundance of the species, it should be noted that these results must be intepreted with caution as the reliability of the effort data is unsure. CPUE for all trawlers is calculated using the tow duration. However this does not account for variations in fishing efficiency or effort due to factors like vessel speed or width of the gear.
The lower map shows the mean CPUE per year over all gear types. Note that for most years, there's only data available from 1 gear type. 

```{r maps}
# CPUE map all years and gears combined
world_map <- borders("worldHires", colour = "gray50", fill = "gray90")

ggplot() +
  world_map +
  geom_point(data = df, aes(x = Longitude, y = Latitude, size = CPUE, color = Gear), alpha = 0.8) +
  coord_fixed(1.3, xlim = c(30, 75), ylim = c(-50, -5)) +  # Adjust these limits to focus on the FAO 51 area
  scale_size_continuous(name = "CPUE") +
  scale_color_brewer(palette = "Set1", name = "Gear Type") +
  ggtitle("CPUE") +
  theme_minimal()

# CPUE spatial scale
# Define longitude and latitude ranges for the grid
lon_range <- seq(30, 80, by = 1)
lat_range <- seq(-5, -50, by = -1)

# Create a grid using expand.grid
grid <- expand.grid(lon = lon_range, lat = lat_range)

# Define function to assign observations to grid cells
assign_to_grid <- function(lon, lat, grid) {
  cell <- rep(NA, nrow(grid))
  for (i in 1:nrow(grid)) {
    cell[i] <- paste0(grid[i, "lon"], "_", grid[i, "lat"])
  }
  
  grid$cell <- cell
  
  cell_assignment <- rep(NA, length(lon))
  for (i in 1:length(lon)) {
    for (j in 1:nrow(grid)) {
      if (lon[i] >= grid[j, "lon"] - 0.5 & lon[i] < grid[j, "lon"] + 0.5 &
          lat[i] >= grid[j, "lat"] - 0.5 & lat[i] < grid[j, "lat"] + 0.5) {
        cell_assignment[i] <- grid[j, "cell"]
        break
      }
    }
  }
  
  return(cell_assignment)
}

# Assign observations to grid cells
df$grid_cell <- assign_to_grid(df$Longitude, df$Latitude, grid)

# Aggregate data by grid cell and calculate total catches (CPUE)
summary_data <- df %>% 
                 group_by(Year, grid_cell) %>% 
                 summarise(sum_cpue = round(sum(CPUE),digits=1),
                           mean_cpue = round(mean(CPUE)),digits=1)

# Split grid cell column into longitude and latitude
summary_data <- transform(summary_data, 
                          lon = as.numeric(sub("_.*", "", grid_cell)),
                          lat = as.numeric(sub(".*_", "", grid_cell)))

for(i in unique(summary_data$Year)){
  
  lfs <- summary_data %>% filter(Year %in% i)

p <- ggplot() +
  world_map +
  geom_tile(data = lfs, aes(x = lon, y = lat, fill = mean_cpue)) +
  coord_fixed(1.3, xlim = c(30, 75), ylim = c(-50, -5)) +
  scale_fill_viridis_c() +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Mean CPUE") +
  ggtitle(paste(i)) +
  theme_minimal()

print(p)

} # End of loop

```


Mean CPUE is calculated per year. These results should be interpreted with caution as for some years, this calculated with CPUEs from different gear types (e.g. gillnets and trawlers). 
```{r cpue}
# Aggregate CPUE over year
cpue2 <- df %>% 
          group_by(Year) %>% 
          summarise(mean_CPUE = mean(CPUE),
                    sd_CPUE = sd(CPUE),
                    cv_CPUE = sd_CPUE/mean_CPUE) 

ggplot(data = cpue2, aes(x = as.factor(Year), y = mean_CPUE, group = 1)) +
  geom_line() +
  theme_minimal() + 
  labs(x = "",
       y = "Mean CPUE")

write.csv(cpue2,"SIOFA_cpue.csv",row.names = F)

```


## IEO dataset
Logbook data containing catches of Portugese dogfish from 2004 to 2022. Years 2006 and 2010 to 2012 are not available. For 2007, only 1 record is available. 
```{r data prep}
# Data preparation
load("../data/inputs/Sharks_logbooks.Rdata") 

# Filter Portugese dogfish, rename some columns 
pdg <- cpue %>% 
        filter(N_Cientifico %in% "Centroscymnus coelolepis") %>% 
        rename(Gear = Arte,
               Species = N_Cientifico,
               Longitude = x3,
               Latitude = y3)

# Check which years/gears are present in the dataset
table(pdg$year,pdg$Gear) 

```

```{r plot CPUE maps}
# CPUE map all years and gears combined
world_map <- borders("worldHires", colour = "gray50", fill = "gray90")

for(i in unique(pdg$year)){
  lfs <- pdg %>% filter(year %in% i)
  
  p <- ggplot() +
        world_map +
        geom_point(data = lfs, aes(x = Longitude, y = Latitude, size = CPUE, color = Gear), alpha = 0.8) +
        coord_fixed(1.3, xlim = c(30, 75), ylim = c(-50, -5)) +  # Adjust these limits to focus on the FAO 51 area
        scale_size_continuous(name = "CPUE") +
        scale_color_brewer(palette = "Set1", name = "Gear Type") +
        ggtitle(paste0("CPUE ",i)) +
        theme_minimal()
  print(p)
} # End of loop

```

For Portugese dogfish, there's only data available for 1 gear type in the different years. Data is available for longlines in 2004 and from 2015-2022. Gillnet CPUEs are available for years 2008-2014
```{r CPUE}
ieo_cpue <- pdg %>% 
             group_by(year) %>%
             summarise(mean_cpue = mean(CPUE),
                       sd_cpue = sd(CPUE),
                       cv_cpue = sd_cpue/mean_cpue)

ggplot(data = ieo_cpue, aes(x = as.factor(year), y = mean_cpue, group = 1)) +
  geom_line() +
  theme_minimal() + 
  labs(x = "",
       y = "Mean CPUE")

write.csv(ieo_cpue, "IEO_cpue.csv",row.names = F)
```


For most years, there’s only data available from 1 gear type in both datasets. Combining CPUEs from multiple gear types could be done using a generalized additive model (GAM) in which the different gear types can be taken into account. However, when there is no data available on catches and/or effort from different gear types within a year, it is not possible to run such a model to derive a reliable estimate of relative abundance. 

```{r CPUE SIOFA and IEO combined plot}
siofa_ieo <- full_join(cpue2 %>% select(c(Year,mean_CPUE)) %>% rename(SIOFA = mean_CPUE),
                       ieo_cpue %>% select(c(year,mean_cpue)) %>% rename(IEO = mean_cpue), by = c("Year"="year")) %>% 
              tidyr::gather(., key = key, value = "Mean_CPUE", SIOFA,IEO)

ggplot(data = siofa_ieo, aes(x = as.factor(Year), y = Mean_CPUE, group = key, color = key)) +
  geom_line() +
  theme_minimal() + 
  labs(x = "",
       y = "Mean CPUE",
       color = "")
```
