---
title: "Annex 6: Priors derived from FishBase"
author: "DELEGATION OF THE EUROPEAN UNION"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
#bibliography: /home/laurie/pCloudDrive/refs.bib
---


# main shark’s species caught by SIOFA

+ Portuguese dogfish (*Centroscymnus coelolepis*) 
+ Birdbeak dogfish (*Deania calcea*) 
+ Kitefin shark (*Dalatias licha*)
+ Gulper sharks (*Centrophorus spp*) 
+ Lanternshark (*Etmopterus spp*)

# fisheries, ,depending on the areas are
  + bottom longline
  + midwater trawl
  + demersal trawl
  + demersal line

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/lh/",
               fig.path  ="../figs/lh-",
               fig.width =10,
               fig.height=8,
               dev       ="png")

iFig=0
iTab=0
```

```{r, eval=FALSE}
install.packages("https://cran.r-project.org/src/contrib/Archive/rfishbase/rfishbase_3.0.0.1.tar.gz", repos=NULL, type='source')

R CMD INSTALL "--no-staged-install" rfishbase
```

```{r figurename, echo=FALSE, out.width = '30%', fig.align="center", eval=FALSE}
setwd("/home/laurie/pCloudDrive/projects/mrag/siofa")
dirRes =file.path(getwd(),"data/results")
dirInp =file.path(getwd(),"data/inputs")
setwd("/home/laurie/pCloudDrive/projects/mrag/siofa/Rmd")

setwd("~/Desktop/active/siofa-haf/Rmd")
setwd("C:/Users/lauri/OneDrive/Desktop/active/siofa-haf/Rmd")
```

```{r}
source("../source/fl_mcupd_fn.R")
```


```{r, pkgs}
library(ggplot2); theme_set(theme_bw())

library(ggpubr)
library(GGally)

library(plyr)
library(dplyr)
library(reshape)

library(FLCore)
library(FLBRP)
library(FLasher)
library(FLife)

library(FLCandy) 
library(mpb)

library(ggplotFL)
library(scales)

library(rfishbase)
library(SPMpriors)
library(FishLife)

library(statcomp)

library(ggcorrplot)
library(ggpubr)

library(R.devices)
```

```{r, eval=TRUE}
library(ggplot2)

# Custom theme for Slidy presentations
theme_slidy_presentation <- function() {
  theme_minimal(base_size = 18) + 
    theme(
      text = element_text(family = "Arial", color = "#333333"),
      plot.title = element_text(face = "bold", size = 24, hjust = 0.5),
      plot.subtitle = element_text(size = 20, hjust = 0.5),
      plot.caption = element_text(size = 16),
      axis.title = element_text(size = 20, face = "bold"),
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      axis.text = element_text(size = 18),
      legend.title = element_text(size = 20, face = "bold"),
      legend.text = element_text(size = 18),
      legend.position = "bottom",
      legend.box = "horizontal",
      panel.grid.major = element_line(color = "#cccccc"),
      panel.grid.minor = element_blank(),
      plot.background = element_rect(fill = "white", color = NA), # Remove plot border
      panel.background = element_rect(fill = "white", color = NA)
    )
}
```

```{r, ggpairs}
my_smooth <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
  geom_point(...,size=.5)+
  geom_smooth(...,method="lm",se=FALSE)}

my_density <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
  geom_density(...,lwd=1)}
```


# Data
```{r, data-fao}
load("../data/inputs/fao_taxa.RData")
```

```{r, data-asfis}
asfis=read.csv("../data/inputs/ASFIS_sp_2021.txt")

names(asfis)[4]="scientificName"
```

```{r, data-fishlife}
spp =transform(rfishbase::fishbase,scientificName=paste(Genus,Species))
```

```{r, list}
sppSiofa=c("Centroscymnus coelolepis","Pailona commun","Portugese dogfish","CYO",
           "Deania calcea","Squale savate","Birdbeak dogfish","DCA",
           "Centrophorus granulosus","Requin chagrin ","ulper shark","GUP",
           "Dalatias licha","Squale liche","Kitefin shark","SCK",
           "Bythaelurus bachi","Requin chat de Bach","Bach's catshark","BZO",
           "Chimaera buccanigella","Chimère bouche-foncée","Dark-mouth chimaera","ZZC",
           "Chimaera didierae","Chimère de Didier","The Falkor chimaera","ZZD",
           "Chimaera willwatchi","Chimère du marin","Seafarer's ghostshark","ZZE",
           "Centroscymnus crepidater","Pailona à long nez","Longnose Velvet Dogfish","CYP",
           "Centroscymnus plunketi","Pailona austral","Plunket shark","CYU",
           "Zameus squamulosus","Squale-grogneur à queue échancrée","Velvet dogfish","SSQ",
           "Etmopterus alphus","Requin lanterne à joues blanches","Whitecheek lanternshark","EZU",
           "Apristurus indicus","Holbiche artouca","Smallbelly catshark","APD",
           "Harriotta raleighana","Chimère à nez rigide","Bentnose rabbitfish","HCR",
           "Bythaelurus tenuicephalus","Requin chat à tête étroite","Narrowhead catshark","BZL",
           "Chlamydoselachus anguineus","Requin lézard","Frilled shark","HXC",
           "Hexanchus nakamurai","Requin griset","Bigeyed six-gill shark","HXN",
           "Etmopterus pusillus","Sagre nain","Smooth lanternshark","ETP",
           "Somniosus antarcticus","Requin dormeur antarctique","Southern sleeper `shark","SON",
           "Mitsukurina owstoni","Requin lutin","Goblin shark","LMO")
sppSiofa=as.data.frame(t(array(sppSiofa,c(4,20))))
sppSiofa=cbind(ldply(strsplit(sppSiofa[,1]," ")),sppSiofa)

names(sppSiofa)=c("Genus","Species","scientificName","french","name","FAO")

sppSiofa=merge(sppSiofa,spp,             by=c("Genus","Species","scientificName"),all.x=T)
sppSiofa=merge(sppSiofa,asfis[,-(12:13)],by=                    "scientificName", all.x=T)

sink(file=NULL)
nulldev()
priors=mdply(sppSiofa[,c("Genus","Species")], function(Genus,Species) {
  
  rtn=try(Plot_taxa(Search_species(Genus=Genus,Species=Species)$match_taxonomy,mfrow=c(3,2)))
  
  if ("try-error"%in%is(rtn)) return(NULL)
  data.frame(trait=names(rtn[[1]][[2]]),data=rtn[[1]][[2]],var=diag(rtn[[1]][[1]]))})
dev.off()
sink()

traits=mdply(sppSiofa[,c("Genus","Species")], function(Genus,Species) {
  
  rtn=try(fl_mcupd(Genus=Genus, Species=Species)$traits)
  
  if ("try-error"%in%is(rtn)) return(NULL)
  rtn})

siofa=merge(priors,traits,by=c("Genus","Species","trait"),all=T)
```

```{r, save-siofa}
save(siofa,file="../data/lh-siofa.RData")
```

```{r, spp}
sppMrag=c(
"Portuguese dogfish"  ,"Centroscymnus coelolepis",	
"Kitefin shark"       ,"Dalatias licha",	
"Birdbeak dogfish"    ,"Deania calcea",												
"Gulper shark"        ,"Centrophorus granulosus",
"Gulper shark"        ,"Centrophorus atromarginatus",	
"Gulper shark"        ,"Centrophorus squamosus",	
"Gulper shark"        ,"Centrophorus harrissoni",	
"Gulper shark"        ,"Centrophorus isodon",	
"Gulper shark"        ,"Centrophorus lesliei",	
"Gulper shark"        ,"Centrophorus lusitanicus",	
"Gulper shark"        ,"Centrophorus moluccensis",	
"Gulper shark"        ,"Centrophorus seychellorum",	
"Gulper shark"        ,"Centrophorus uyato",	
"Gulper shark"        ,"Centrophorus westraliensis",	
"Gulper shark"        ,"Centrophorus zeehaani",	
"Lantern shark"       ,"Etmopterus alphus",	
"Lantern shark"       ,"Etmopterus bigelowi",	
"Lantern shark"       ,"Etmopterus brosei",	
"Lantern shark"       ,"Etmopterus compagnoi",	
"Lantern shark"       ,"Etmopterus evansi",	
"Lantern shark"       ,"Etmopterus fusus",	
"Lantern shark"       ,"Etmopterus gracilispinis",	
"Lantern shark"       ,"Etmopterus granulosus",	
"Lantern shark"       ,"Etmopterus pusillus",	
"Lantern shark"       ,"Etmopterus sentosus")

sppMrag=data.frame(name=sppMrag[seq(1,length(sppMrag),2)],scientificName=sppMrag[seq(2,length(sppMrag),2)])

spp =transform(rfishbase::fishbase,scientificName=paste(Genus,Species))
sppMrag=merge(sppMrag,spp,             by="scientificName",all.x=T)
sppMrag=merge(sppMrag,asfis[,-(12:13)],by="scientificName",all.x=T)

sink(file=NULL)
nulldev()
priors=mdply(sppMrag[,c("Genus","Species")], function(Genus,Species) {
  
  rtn=try(Plot_taxa(Search_species(Genus=Genus, Species=Species)$match_taxonomy,mfrow=c(3,2)))
  
  if ("try-error"%in%is(rtn)) return(NULL)
  data.frame(trait=names(rtn[[1]][[2]]),data=rtn[[1]][[2]],var=diag(rtn[[1]][[1]]))})
dev.off()
sink()

traits=mdply(sppMrag[,c("Genus","Species")], function(Genus,Species) {
  
  rtn=try(fl_mcupd(Genus=Genus, Species=Species)$traits)
  
  if ("try-error"%in%is(rtn)) return(NULL)
  rtn})

mrag=merge(priors,traits,by=c("Genus","Species","trait"),all=T)
```

```{r, save-mrag}
save(mrag,file="../data/mrag.RData")
```

```{r, cor}
lh=cast(siofa,Genus+Species~trait,value="data") 
save(lh,file="../data/lh.RData")

pr1=ggpairs(transform(lh[,c("Loo","K","Lm","G","M","r","h")],'l50:loo'=Lm/Loo)[c(1:3,7,8,4:6)])+
              #mapping = ggplot2::aes(color=rank(dat$hmsy)<6))+
   theme_bw()+theme(legend.position = "none", 
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA", angle=90), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"),
                    
                    axis.title.x = element_blank(), 
                    axis.text.x  = element_blank(), 
                    axis.ticks.x = element_blank(), 
                    
                    plot.margin = unit(c(0, 0, 0, 1), "lines"),
                    panel.background = element_rect(fill   ="NA", colour="NA"), 
                    #panel.border     = element_rect(fill   ="NA", colour="NA"), 
                    panel.grid.major = element_line(colour ="NA"), 
                    panel.grid.minor = element_line(colour ="NA"))
pr1
```


```{r, m, fig.height=8,fig.width=8}
m=ddply(subset(siofa,trait=="M"), .(Genus,Species,trait), with,
           data.frame(iter=seq(100),dev=exp(rnorm(100,data,var)),mean=data,sd=var^0.5))
r=ddply(subset(siofa,trait=="r"), .(Genus,Species,trait), with,
           data.frame(iter=seq(100),dev=rnorm(100,data,var),mean=data,sd=var^0.5))
rfpts=rbind(cbind(Derivation="M",mutate(m,Fmsy=0.41*dev,Flim=1.5*Fmsy,Fcrash=2*Fmsy)),
            cbind(Derivation="r",mutate(r,Fmsy=0.5*dev, Flim=1.5*Fmsy,Fcrash=2*Fmsy))) 

ggdensity(m,x="dev",fill="red")+
  facet_wrap(paste(Genus,Species,sep="\n")~.)+
  scale_x_continuous(limits=c(0,.3))+
  theme_slidy_presentation()+    
  theme(legend.position = "none", 
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA"), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"))
```

```{r, r, fig.height=8,fig.width=8}
ggdensity(r,x="dev",fill="red")+
  facet_wrap(paste(Genus,Species,sep="\n")~.)+
  scale_x_continuous(limits=c(0,.1), breaks=c(0.01,0.05,0.09))+
  xlab("Population Growth Rate (r)")+
    theme_slidy_presentation()+
    theme(legend.position = "none", 
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA"), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"))
```

**Figure `r iFig=iFig+1; iFig`** Estimates of population growth $r$ based on life-history theory


```{r, fmsy, fig.height=12,fig.width=10}
ggdensity(rfpts,x="Fmsy",fill="Derivation")+
  facet_wrap(paste(Genus,Species,sep="\n")~.,scale="free", ncol=3)+
  scale_x_continuous(limits=c(0,0.3),breaks=c(0.1, 0.2, 0.3))+
  xlab(expression(F[MSY]))+
    theme_slidy_presentation()+
    theme(legend.position = "bottom", 
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA"), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"))
```

**Figure `r iFig=iFig+1; iFig`** Estimates of $F_{MSY}$ based on life-history theory

# MRAG

```{r, cor2}
dat=cast(mrag,Genus+Species~trait,value="data")

pr1=ggpairs(transform(subset(dat,Genus%in%c("Centrophorus","Etmopterus"))[,c("Genus","Loo","K","Lm","G","M","r","h")],
                      'l50:loo'=Lm/Loo)[c(1:3,7,8,4:6)],
              mapping=ggplot2::aes(color=Genus))+
   theme_slidy_presentation()+    theme(legend.position = "none", 
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA", angle=90), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"),
                    
                    axis.title.x = element_blank(), 
                    axis.text.x  = element_blank(), 
                    axis.ticks.x = element_blank(), 
                    
                    plot.margin = unit(c(0, 0, 0, 1), "lines"),
                    panel.background = element_rect(fill   ="NA", colour="NA"), 
                    #panel.border     = element_rect(fill   ="NA", colour="NA"), 
                    panel.grid.major = element_line(colour ="NA"), 
                    panel.grid.minor = element_line(colour ="NA"))
pr1
```


```{r, m1, fig.height=10,fig.width=8}
m=ddply(subset(mrag,trait=="M"), .(Genus,Species,trait), with,
           data.frame(iter=seq(100),dev=exp(rnorm(100,data,var)),mean=data,sd=var^0.5))
r=ddply(subset(mrag,trait=="r"), .(Genus,Species,trait), with,
           data.frame(iter=seq(100),dev=rnorm(100,data,var),mean=data,sd=var^0.5))
rfpts=rbind(cbind(Derivation="M",mutate(m,Fmsy=0.41*dev,Flim=1.5*Fmsy,Fcrash=2*Fmsy)),
            cbind(Derivation="r",mutate(r,Fmsy=0.5*dev, Flim=1.5*Fmsy,Fcrash=2*Fmsy)))
```


```{r, m2, fig.height=4,fig.width=8}
ggdensity(m,x="dev",fill="Species",alpha=0.25)+
  facet_wrap(Genus~.)+
  scale_x_continuous(limits=c(0,.3))+ 
  xlab("M")+
  theme_slidy_presentation()+    
              theme(legend.position = "none", 
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA"), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"))
```

**Figure `r iFig=iFig+1; iFig`** Estimates of natural mortality based on life-history theory


```{r, r1, fig.height=4,fig.width=8}
ggdensity(r,x="dev",fill="Species",alpha=0.25)+
  facet_wrap(Genus~.)+
  xlab("Population Growth Rate (r)")+
  scale_x_continuous(limits=c(0,.1),breaks=c(0.05,0.1))+
     theme_slidy_presentation()+    
              theme(legend.position = "none", 
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA"), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"))
```

**Figure `r iFig=iFig+1; iFig`** Estimates of population growth $r$ based on life-history theory

```{r, fmsy1, fig.height=12,fig.width=10}
ggdensity(rfpts,x="Fmsy",fill="Derivation")+
  facet_wrap(paste(Genus,Species,sep="\n")~.,scale="free")+
  scale_x_continuous(limits=c(0,0.2), breaks=c(0, 0.1, 0.2))+
  theme_slidy_presentation()+
              theme(legend.position = "bottom",   
                    axis.title.y = element_text(colour='NA'), 
                    axis.text.y  = element_text(colour="NA"), 
                    axis.ticks.y = element_line(colour="NA"),
                    axis.ticks =   element_line(colour="NA"))
```

**Figure `r iFig=iFig+1; iFig`** Estimates of $F_{MSY}$ based on life-history theory


```{r, eval=FALSE}
library(ggplot2)  
library(RColorBrewer)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(lubridate)
library(plyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(gridExtra)

library(aquamapsdata)
library(raquamaps)
library(gfwr)

library(leaflet)
library(raster)
library(aquamapsdata)

library(aquamapsdata)
#download_db(force = TRUE)
default_db("sqlite")

ss =read_sf(file.path("/home/laurie/pCloudDrive/shapeFiles/SS_Boundary/Sargasso_Sea_Geographical_Area_of_Collaboration.shp"))
ssE=read.csv("/home/laurie/Dropbox/Sargasso FFEM Shared 2021/06 Deliverables/analysis/inputs/gfw/bq-results-20221012-060605-1665555242269.csv")

world <- ne_countries(scale = "medium", returnclass = "sf")

#d2p <- nd %>% 
#  group_by(lat, lon, gear) %>% 
#  summarise(fishing_time = sum(fishing_time))

ss =read_sf(file.path("/home/laurie/pCloudDrive/shapeFiles/SS_Boundary/Sargasso_Sea_Geographical_Area_of_Collaboration.shp"))
ssE=read.csv("/home/laurie/Dropbox/Sargasso FFEM Shared 2021/06 Deliverables/analysis/inputs/gfw/bq-results-20221012-060605-1665555242269.csv")

world <- ne_countries(scale = "medium", returnclass = "sf")

iccat=read_sf("/home/laurie/pCloudDrive/shapeFiles/ICCAT/ICCAT_draft_ecoregions_May 2022.shp")

ggplot(subset(ssE,geartype=="trawlers")) +
    geom_tile(aes(x=lon_bin_center, y=lat_bin_center, fill=log10(fishing_hours))) +
    scale_fill_distiller(palette = "Blues", direction = 1) +
    geom_sf(data = ss, fill = NA)+
    #geom_s f(data = iccat, fill = NA, col="red")+
    geom_sf(data = world) +
  theme_bw()+
  labs(x = "Longitude", y= "Latitude") +
      coord_sf(expand = FALSE,
             xlim=c(c(-40,-80)), ## set the plot longitude extents
             ylim=c(20,40)) ## set the plot latitude extents

key <- am_search_fuzzy("Illex illecebrosus")$key
ras <- am_raster(key)

# show the native habitat map
am_map_leaflet(ras, title = "Shortfin Squid") %>%
leaflet::fitBounds(lng1 =-75, lat1=10, lng2=-30, lat2=55) %>%
  addPolygons(data=ss, weight = 2)
```

