---
title: "**SIOFA Sharks**"
subtitle: "Centroscymnus coelolepis"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
fig_width: 6 
fig_height: 4 
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
#bibliography: /home/laurie/pCloudDrive/refs.bib
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path="cache/pdog/",
               fig.path  ="../figs/pdog/",
               fig.width =12,
               fig.height=8,
               dev       ="png")
iFig=0
```

```{r setup}
library(openxlsx)  # For reading Excel files
library(lubridate) # For handling date and time
library(plyr)      # For data manipulation
library(dplyr)     # For data manipulation
library(reshape)   # For data manipulation
library(ggplot2)   # For plotting
library(ggpubr)    # For plotting
library(mapdata)   # For map-based data
library(viridis)   # For color scales, better for visibility
```

```{r, setwd}
setwd("/home/laurie/Desktop/projects/siofa/Rmd")
```

```{r, data}
# Data Preparation
# Load data from an Excel file into data frames
cpue = read.xlsx("../data/inputs/SIOFA_DWS-2023-01-Data.xlsx", "dws_data_catch", 
                 startRow = 1, detectDates = TRUE, skipEmptyRows = TRUE)
biol = read.xlsx("../data/inputs/SIOFA_DWS-2023-01-Data.xlsx", "dws_data_bio_sampling",
                 startRow = 1, detectDates = TRUE, skipEmptyRows = TRUE)

# Adjusting time format
adjust_time_format<-function(time)
  format(as.POSIXct(time * 86400, origin = "1970-01-01", tz = "UTC"), "%H:%M:%S")

cpue$SettingTime=adjust_time_format(cpue$SettingTime)
cpue$HaulingTime=adjust_time_format(cpue$HaulingTime)
cpue$Time       =adjust_time_format(cpue$Time)
cpue$Year       =year(cpue$Date)

cpue=subset(cpue,speciesScientificName=="Centroscymnus coelolepis")
biol=subset(biol,speciesScientificName=="Centroscymnus coelolepis")

catchSmry<-function(data) {
  # Filter out NAs
  filtered=filter(data, !is.na(Latitude), !is.na(Longitude), !is.na(Year))
  # Group data
  rtn=group_by(filtered, speciesScientificName, Gear, Year, Latitude, Longitude)
  
  rtn$Species=gsub(" ","\n",rtn$speciesScientificName)
  
  rtn}

cSmry=catchSmry(cpue)
```

```{r, theme}
theme_my<-function(size=20) {
  theme_minimal() +
    theme(text = element_text(size=size),
          plot.title = element_text(face = "bold", size = size+2),
          axis.title = element_text(size = size-1),
          legend.position = "bottom",
          legend.title = element_blank(),
          panel.grid.major = element_line(color = "gray80"),
          panel.grid.minor = element_blank())}
```

### Catch Weight by Gear and Year Map



### Species Distribution by Length

```{r species-distribution-length}
biola <- subset(biol, select = c("OPE_ID", "MinLength","MaxLength","NbFemale","NbMale",
                               "NbUndetermined","Length", "Sex"))
df3   <- merge(cpue, biola, by = "OPE_ID")

ggplot(subset(df3, !is.na(Length)), aes(x = Length)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  labs(title = "Distribution of Fish Sizes", x = "Length", y = "Count")
```

**Figure `r iFig=iFig+1; iFig`** Species Distribution by Length

### Gender Distribution

```{r gender-distribution}
table(df3$speciesScientificName, df3$Sex, useNA="always")
df3$Sex[df3$Sex=="F "]   <- "F"

# Aggregating data by species and gender
gender_distribution <- df3 %>%
  group_by(speciesScientificName, Sex) %>%
  summarize(Count = n(), .groups = 'drop')

ggplot(gender_distribution, aes(x = Sex, y = Count, fill = Sex)) +
  geom_bar(stat = "identity") +
  theme_my() +
  labs(title = "Gender Distribution by Species", x = "Gender", y = "Count") +
  scale_fill_brewer(palette = "Spectral", name = "Gender")
```

**Figure `r iFig=iFig+1; iFig`** Gender Distribution

### Catch Weight Over Time

```{r catch-weight-over-time}
catch_summary <- cpue %>%
  group_by(Year, speciesScientificName) %>%
  summarize(TotalCatchWeight = sum(catchWeight, na.rm = TRUE), .groups = 'drop')

ggplot(catch_summary, aes(x = Year, y = TotalCatchWeight, color = speciesScientificName)) +
  geom_line() +
  theme_my() +
  labs(title = "Catch Weight by Species Over Time", x = "Year", y = "Total Catch Weight") +
  theme(legend.position = "bottom")
```

**Figure `r iFig=iFig+1; iFig`** Catch Weight Over Time

### Overall Catch Weight Over Time

```{r overall-catch-weight-over-time}
overall_catch <- catch_summary %>%
  group_by(Year) %>%
  summarize(TotalCatchWeight = sum(TotalCatchWeight, na.rm = TRUE), .groups = 'drop')

ggplot(overall_catch, aes(x = Year, y = TotalCatchWeight)) +
  geom_line() +
  theme_my() +
  labs(title = "Overall Catch Weight Over Time", x = "Year", y = "Total Catch Weight")
```

**Figure `r iFig=iFig+1; iFig`** Overall Catch Weight Over Time

```{r, map1}
summarized_catch <- cpue %>%
  filter(!is.na(Latitude) & !is.na(Longitude) & !is.na(Year)) %>%
  group_by(speciesScientificName, Gear, Year, Latitude, Longitude) %>%
  summarize(TotalCatch = sum(catchWeight, na.rm = TRUE))

world_map <- borders("worldHires", colour = "gray50", fill = "gray90")

ggplot(ddply(summarized_catch,.(Longitude,Latitude,Gear), with, 
             data.frame(data=sum(TotalCatch))))+
        world_map +
        geom_point(aes(x=Longitude, y=Latitude, size=data), alpha=0.8) +
        coord_fixed(1.3, xlim = c(30, 100), ylim = c(-50, 5)) +
        theme_my()+
  facet_wrap(~Gear)
```

**Figure `r iFig=iFig+1; iFig`** 

